### 2019-03-23 paper/sex_selection_spacing-ver2.tex
Edit distributional results section for grammar

The results above show consistent increases in average spacing across the board.
Average spacing is a convenient way to understand the overall changes in behavior but may 
hide important differences in the distribution of spacing.
Specifically, whether these increases in average spacing come from fewer births after 
very short intervals, because of even longer spacing from births with already long 
spacing, or a general increase in spacing have important implications for what effects the
increase in average spacing are likely to have and how to interpret the motivation for
the changes.
On one hand, if there still are many births that occur after very short intervals, the 
increase in average spacing is unlikely to have a substantial impact on health outcomes
since the literature on birth spacing and child health suggests that an interpregnancy 
interval of less than 18 months is associated with a higher risk of adverse health 
outcomes and mortality \citep{Conde-Agudelo2012}.
On the other hand, we may see a reduction in early observed births and more long spells if 
families with fewer sons are both more likely to have a very early pregnancy and to 
continue through pregnancies and sex-selective abortions until they conceive a son.%
\footnote{
It is unclear what the optimal response in timing of pregnancies to the availability of 
sex selection should be.
The expected birth intervals become longer with the use of sex selection, which may be 
costly for parents leading them to try to conceive even earlier than previously.
Conversely, if parents know that short birth intervals are detrimental to the next child's
health, and with sex selection the next child is likely be a boy, it might make sense to 
wait longer before conceiving.
}

@@

Instead of averaging across the entire sample, I calculate the conditional survival curves 
for an average woman using the method detailed in Section \ref{sec:strategy}.
For each combination of education and spell, I use values based on the average age at the 
start of the spell.
Furthermore, I use the majority categories for the categorical explanatory  variables, 
which means no ownership of land and not in a scheduled caste or tribe.
The characteristics used do not change across the four decades to ensure that composition 
effects do not drive any changes.
To complement this approach, Appendix Tables \ref{tab:p25_p50_p75_low}, 
\ref{tab:p25_p50_p75_med}, and \ref{tab:p25_p50_p75_high} show 25th, 50th, and 75th 
percentiles durations with bootstrapped standard errors based on averages across all women 
in the each subsample as for the results above.

In the interest of brevity, I discuss only two subsamples.
Figures \ref{fig:pps_low} and \ref{fig:pps_high} show spacing across sex compositions for 
the second, third, and fourth spells for the first period, 1972--1984, and the last 
period, 2005--2016, for rural women with no education and for urban women with eight or 
more years of education.
The Appendix shows conditional survival curves for all groups and decades.

@@

For the third spell the son preference driven short spacing is evident throughout the
distribution and becomes more pronounced over the four decades.
Families with two girls are especially likely to have a birth early with 25 percent
having a child within 12 months of the beginning of the spell (21 months after the prior
second birth).
That number is only slightly smaller four decades later. 
The fourth spell shows changes in the distribution that appear to be more pronounced
than for the third spell, although these results are based on fewer observed births.
For families with three girls, 75 percent have had their fourth child within 36 months 
after the beginning of the spell, but this occurs more than a year later for families
with two or more boys.

@@

For the third spell, the 1972--1984 period show a close to standard pattern throughout, 
although, as mentioned above, the sex ratios for women without girls was substantially 
higher than the natural sex ratio either because of recall error or early use of 
sex selection.
The correlation between the increased use of sex selection and the reversal of the 
standard spacing pattern show up clearly in the 2005--2016 period for the third spell with 
spacing to the third child consistently longer with two girls than with either two boys 
or one boy and one girl throughout the entire spell.
This pattern is most consistent with early and continued use of sex selection, rather 
than a general reduction in the likelihood of very short birth intervals.
It also means, that previous girls can expect a longer period without a younger sibling,
which may reduce the sibling competition for resources.

There is an even more pronounced rightward and upward shift for spacing after only
girls in the fourth spell for the first years of the spell.%
\footnote{
The caveat is that there are few births to base the results on,
especially for women with two or more sons.
} 
Furthermore, the differences are substantial across most of the distribution.
Twelve months after the beginning of the spell there is a more than ten 
percentage points difference in the conditional survival curves for only girls
and only boys.
Both of the spells therefore show changes in the distribution of births that is in line
with extensive use of sex selection among families with no or one son.

>>>>

The results above show consistent increases in average spacing across the board.
Average spacing is a convenient way to understand the overall changes in behavior but may 
hide important differences in the distribution of spacing.
Specifically, whether these increases in average spacing come from fewer births after 
very short intervals, because of even longer spacing from births with already long 
spacing, or a general increase in spacing have important implications for what effects the
increase in average spacing are likely to have and how to interpret the motivation for
the changes.
On one hand, if there still are many births that occur after very short intervals, the 
increase in average spacing is unlikely to have a substantial impact on health outcomes
since the literature on birth spacing and child health suggests that an interpregnancy 
interval of fewer than 18 months is associated with a higher risk of adverse health 
outcomes and mortality \citep{Conde-Agudelo2012}.
On the other hand, we may see a reduction in early observed births and more long spells if 
families with fewer sons are both more likely to have a very early pregnancy and to 
continue through pregnancies and sex-selective abortions until they conceive a son.%
\footnote{
It is unclear what the optimal response in the timing of pregnancies to the availability 
of sex selection should be.
The expected birth intervals become longer with the use of sex selection, which may be 
costly for parents leading them to try to conceive even earlier than previously.
Conversely, if parents know that short birth intervals are detrimental to the next child's
health, and with sex selection, the next child is likely a boy, it might make sense to 
wait longer before conceiving.
}

@@

Instead of averaging across the entire sample, I calculate the conditional survival curves 
for an average woman using the method detailed in Section \ref{sec:strategy}.
For each combination of education and spell, I use values based on the average age at the 
start of the spell.
Furthermore, I use the majority categories for the categorical explanatory  variables, 
which means no ownership of land and not in a scheduled caste or tribe.
The characteristics used do not change across the four decades to ensure that composition 
effects do not drive any changes.
To complement this approach, Appendix Tables \ref{tab:p25_p50_p75_low}, 
\ref{tab:p25_p50_p75_med}, and \ref{tab:p25_p50_p75_high} show 25th, 50th, and 75th 
percentiles durations with bootstrapped standard errors based on averages across all women 
in each subsample as for the results above.

In the interest of brevity, I discuss only two subsamples.
Figures \ref{fig:pps_low} and \ref{fig:pps_high} show spacing across sex compositions for 
the second, third, and fourth spells for the first period, 1972--1984, and the last 
period, 2005--2016, for rural women with no education and urban women with eight or 
more years of education.
The Appendix shows conditional survival curves for all groups and decades.

@@

For the third spell, the son preference driven short spacing is evident throughout the
distribution and becomes more pronounced over the four decades.
Families with two girls are especially likely to have a birth early with 25 percent
having a child within 12 months of the beginning of the spell (21 months after the prior
second birth).
That number is only slightly smaller four decades later. 
The fourth spell shows changes in the distribution that appear to be more pronounced
than for the third spell, although these results are based on fewer observed births.
For families with three girls, 75 percent have had their fourth child within 36 months 
after the beginning of the spell, but this occurs more than a year later for families
with two or more boys.

@@

For the third spell, the 1972--1984 period show a close to standard pattern throughout, 
although, as mentioned above, the sex ratios for women without girls was substantially 
higher than the natural sex ratio either because of recall error or early use of 
sex selection.
The correlation between the increased use of sex selection and the reversal of the 
standard spacing pattern show up clearly in the 2005--2016 period for the third spell with 
spacing to the third child consistently longer with two girls than with either two boys 
or one boy and one girl throughout the entire spell.
This pattern is most consistent with early and continued use of sex selection, rather 
than a general reduction in the likelihood of very short birth intervals.
It also means, that previous girls can expect more time without a younger sibling,
which may reduce the sibling competition for resources.

There is an even more pronounced rightward and upward shift for spacing after only
girls in the fourth spell for the first years of the spell.%
\footnote{
The caveat is that there are few births to base the results on,
especially for women with two or more sons.
} 
Furthermore, the differences are substantial across most of the distribution.
Twelve months after the beginning of the spell there is a more than ten 
percentage points difference in the conditional survival curves for only girls
and only boys.
Both of the spells, therefore, show changes in the distribution of births that is in line
with extensive use of sex selection among families with no or one son.
